---
title: "Ticino Alpine Transport Accessibility Report"
author: "group 4"
date: "19/06/2026"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(sf)
library(knitr)
```

# Overview

This report describes the transport accessibility work completed for our
Ticino region. The project is organised as an R package. The files in
the package read station data, query public transport routes, cache API
results, calculate waiting times, and visualise the results on a regional map.

The work done in the project follows this process:

1. Store the cleaned station table in `data/saved_data.rda`.
2. Store the map geometry in `data/ch.rda`.
3. Read station data with `read_csv_data()`.
4. Query routes with `get_routes()` and `get_routes_all()`.
5. Reuse saved API responses from the `cache/` folder.
6. Calculate waiting-time indicators with `calculate_waiting()`.
7. Plot waiting-time and route maps with the plotting functions.

# Station Data

The station data is stored in `data/saved_data.rda` as `saved_data`. The
package function `read_csv_data()` returns this object.

```{r station-data}
stations <- read_csv_data()
stations$station_id <- as.character(stations$station_id)
stations$is_origin <- as.logical(stations$is_origin)

kable(
  stations,
  caption = "Station data used for our Ticino / Alpine region"
)
```

The dataset contains `r nrow(stations)` stations. One station is marked as the
origin and the remaining stations are destinations.

```{r origin-destination-summary}
origin <- stations[stations$is_origin, ]
destinations <- stations[!stations$is_origin, ]

kable(
  origin[, c("city", "station_name", "station_id", "latitude", "longitude")],
  caption = "Origin station"
)

kable(
  destinations[, c("city", "station_name", "station_id", "population")],
  caption = "Destination stations"
)
```


# Waiting-Time Calculation

The function `calculate_waiting()` in `R/read_stations.R` calculates waiting
times by:

1. Reading the station data.
2. Finding the origin station.
3. Calling `get_routes_all()` for each query time.
4. Comparing each requested time with the available departure times.
5. Keeping the smallest non-negative wait for each destination and query time.
6. Summarising the results with median waiting time, mean waiting time, and
   number of queries.

The available cache files match the query date `06/20/2026` and the query
times `08:00`, `10:00`, `12:00`, and `14:00`, so the calculation below uses
the saved route results.

```{r waiting-calculation}
waiting <- calculate_waiting(
  date = "06/20/2026",
  query_times = c("08:00", "10:00", "12:00"),
  num = 3
)

kable(
  waiting[, c("city", "station_name", "population", "median_wait", "mean_wait", "n_queries")],
  caption = "Waiting-time indicators calculated by calculate_waiting()"
)
```

# Waiting-Time Map

The waiting-time map is produced by `plot_waiting_map()`. 

In the plot:
- the black star is the origin;
- point size represents population;
- point colour represents median waiting time in minutes.

```{r waiting-map, fig.width=8, fig.height=6}
plot_waiting_map(
     date = "06/22/2026",
     query_times = c("08:00", "10:00", "12:00"),
     num = 3
)
```

# Route Map

This file contains one route from the origin station to each destination for
the 17:30 query time. The function `plot_route_map()` reads the nested route
points and draws route lines on the same `ch` map geometry.

```{r route-map, fig.width=8, fig.height=6}
plot_route_map(8505204, "06/22/2026", "6:56")
```

# Single-Route Plot

The file `R/plot_route.R` contains `plot_route()`, which plots one selected
route. The example below uses the first destination route from the cached
17:30 all-destination result.

```{r single-route-map, fig.width=8, fig.height=6}
plot_route(8505400, 8505300, "06/22/2026", "10:56", 3)
```
